### 2018-12-27 papers/journal/encoding.tex
Improvements in the reachability section edited slightly
intro addition

\subsection{The key improvement}
The key improvement of SMT tool over the BMC is the
encoding of reachability, which was done using an enumeration of paths. 
%
In SMT tool, reachability is encoded in two different
ways in constraints~\eqref{eq:reach1} and~\eqref{eq:drop3}.
%
The reachability is recursively defined in~\eqref{eq:drop3} and has
trivial solutions by making all $r'$s true.
%
However, the trivial solutions are disallowed by constraint~\eqref{eq:drop4} and we find
only the solutions that captures the evidence of unreachability.
%
On the other hand,
we have added length of paths in our reachability encoding in constraint~\eqref{eq:reach1},
which needs relatively more auxiliary variables.
%
This is because constraint~\eqref{eq:reach2} has only positive
occurrences of the reachability variables and if we had defined
$r$s in ~\eqref{eq:reach1} without paths,
the circular dependencies in the recursive definitions of $r$s
may have resulted in spurious satisfying assignments that
do not encode reachability.
%
By adding the path length, we break the circular dependencies, the
constraint remains polynomial in size, and satisfying assignments only
correspond to correct reachability.

@@

We have encoded the following variants of VTSs.
%
The variants are due to different combinations of constraints on the
activity of molecules on the nodes and edges.
%
\begin{enumerate}
\item Every present molecule is considered to be active.
\item Activity of molecules on the nodes is based on Boolean function of presence of other molecules. 
\item Activity of molecules on the edges is based on Boolean function of presence of other molecules
\item Activity of molecules both on the edges and nodes is based on Boolean function of presence of other molecules.
\item Activity of molecules on the edges is driven by pairing inhibition.
\item Activity of molecules on the nodes is based on Boolean function of presence of other molecules and on edge by pairing inhibition.
\end{enumerate}
%
% For the version A, every present molecule is in active state.
%
In the table~\ref{tab:grph}, we present the constraints involved in each version.
%
The column two of the table shows that the constraints V1-V8, R1-R2, and D1-D4 are present
for every variant.
%
The column three of the table lists the constraints that are different among the variants.
%
%
One of the restriction is where the activity of the present molecule is dependent on the presence of the other molecules.
%
For example in version B, D, F activity of the molecules on the node is a boolean function of the presence of other molecules on that node; Anb.
%
Similarly for the case of the edge in version C and D; Aeb.
%
In the case of versions F the activity of the molecules on the edges
is described by pairing matrix; Aep.

@@

\input{qbf-table}
%
In the table~\ref{tab:satqbf-graph}, you can see that SAT version was not able to scale it to the desired results.
%
This is due to the fact that directly translating a quantified formula into a quantifier-free  by introducing a vector of variables for each boolean possibility will blow up the size of the formula and may not be solvable by a SAT solver.

>>>>

\subsection{The key improvement}
The key improvement of SMT tool over the BMC is the
encoding of reachability.
% which was done using an enumeration of paths. 
%
In SMT tool, reachability is encoded in two different
ways in constraints~\eqref{eq:reach1} and~\eqref{eq:drop3}.
%
The reachability is recursively defined in~\eqref{eq:drop3} and has
trivial solutions by making all $r'$s true.
%
However, the trivial solutions are disallowed by constraint~\eqref{eq:drop4} and we find
only the solutions that captures the evidence of unreachability.
%
On the other hand,
we have added length of paths in our reachability encoding in constraint~\eqref{eq:reach1},
which needs relatively more auxiliary variables.
%
This is because constraint~\eqref{eq:reach2} has only positive
occurrences of the reachability variables and if we had defined
$r$s in ~\eqref{eq:reach1} without paths,
the circular dependencies in the recursive definitions of $r$s
may have resulted in spurious satisfying assignments that
do not encode reachability.
%
By adding the path length, we break the circular dependencies, the
constraint remains polynomial in size, and satisfying assignments only
correspond to correct reachability.

\input{variation-table}

@@

We have encoded the following variants of VTSs.
%
The variants are due to different combinations of constraints on the
activity of molecules on the nodes and edges.
%
\begin{enumerate}
\item Every present molecule is considered to be active.
\item Activity of molecules on the nodes is based on Boolean function of presence of other molecules. 
\item Activity of molecules on the edges is based on Boolean function of presence of other molecules
\item Activity of molecules both on the edges and nodes is based on Boolean function of presence of other molecules.
\item Activity of molecules on the edges is driven by pairing inhibition.
\item Activity of molecules on the nodes is based on Boolean function of presence of other molecules and on edge by pairing inhibition.
\end{enumerate}
%
% For the version A, every present molecule is in active state.
%

In the table~\ref{tab:var-grph}, we present the constraints involved in each version.
%
The column two of the table shows that the constraints V1-V8, R1-R2, and D1-D4 are present
for every variant.
%
The column three of the table lists the constraints that are different among the variants.
%
%
One of the restriction is where the activity of the present molecule is dependent on the presence of the other molecules.
%
For example in version B, D, F activity of the molecules on the node is a boolean function of the presence of other molecules on that node; Anb.
%
Similarly for the case of the edge in version C and D; Aeb.
%
In the case of versions F the activity of the molecules on the edges
is described by pairing matrix; Aep.

@@

\input{satqbf-table}
%
In the table~\ref{tab:satqbf-graph}, you can see that SAT version was not able to scale it to the desired results.
%
This can be due to the fact that directly translating a quantified formula into a quantifier-free  by introducing a vector of variables for each boolean possibility will blow up the size of the formula and may not be solvable by a SAT solver.

### 2018-12-27 papers/journal/experiments.tex
intro addition

%We have developed four different tools (\sattool, \smttool, \qbftool, \ourtool) for implementation of proposed encodings.
%
We have developed the {\vtstool} tool to conjecture the properties of the VTS. 
%
The toolchain is divided into three different blocks. 
%
Checking existential condition, checking the universal condition and performing synthesis on the incomplete VTS. 
%
The existential condition is subdivided into two sub-block, using bounded model checking and using constraint solving. 
%
We will refer to existential subblocks as {\sattool} and {\smttool}, universal block as {\qbftool} and synthesis block as {\ourtool}.

\subsection{Implementation}
{\sattool} is based on the architecture of a Bounded model checker (CBMC) and uses an enumeration of paths for encoding the reachability problem.
%
{\smttool} is based on the SMT solver Z3 and uses the efficient encoding of finding a least fixed point to solve reachability problem. 
%
{\qbftool} is based on the QBF encoding of the universal condition and uses {\depqbf} solver to solve the generated QBF. 
%
The {\ourtool} is a synthesis tool for synthesizing incomplete information in the input VTS.
%
All the experiments were done on a machine with  Intel(R) Core(TM) i7-2600 CPU @ 3.40GHz processor and 16GB RAM with 30 min (1800 sec) timeout.

@@

The CBMC architecture consists of a C-language front-end (CFE) and a SAT Solver back-end (SAT). 
%
CBMC front end accepts ANSI-C programs with some special annotations (called assumes and assertions) to express constraints on the model and properties to be checked. 
%		
In our experiments, the vesicle transport network was modelled as a non-deterministic C-program manipulating a graph with labelled edges with the fusion rules and steady state properties expressed as constraints using assumes. 
%		
The correspondence relationship between connectedness conditions
and guarantees for steady state were expressed as assertions to be checked. 
%		
Given an exploration depth provided by the user, which in our case corresponds to the size of the graph, CBMC verifies the properties by executing the following steps:
	\begin{itemize}
	\item Convert (using CFE) the model and the properties into a Boolean formula (verification condition) such that the property is true of all behaviors of the model up to the specified
depth iff the Boolean formula is valid.
	\item Check (using a SAT solver) validity of the verification condition. CBMC reports successful verification if the formula is found to be valid. If not, then CBMC produces a counterexample,
an assignment to the variables of the model, that is a witness to the violation of the property.
	\end{itemize}

@@

\textbf{{\smttool} (SMT based tool)}:
	We have implemented the encodings for each variants using the python interface of Z3 in a tool {\smttool}. 
%
Our tool allows the user to choose a model and the size
of the network besides other parameters like connectivity and number of parallel edges. 
%		
It uses Z3 Python interface to build the needed constraints and applies Z3 solver on the constraints to find a model ( a satisfying assignment that respects the constraints). 
%
This tool also translates the satisfying model found by Z3 into
a VTS and presents a visual output to the user in form of annotated graph. 
%
We also visually report the dropped edges required to disconnect the graph, it gives information about the connectivity of the graph.
%
To illustrate usability of our tool in the last column of the table 1, we present the minimum connectedness needed for the different variants after applying our tool for sizes from 2 to 10. 
%
We found no graph for the variant A with constraints Ann and Aen. 
%
Replacing constraint on the node of Ann with Anb (variant B)
does not affect the outcome. 
%
If we allow every present molecule to stay active
(Ann) but constraint the edge by a boolean function (Aeb) the resultant VTS has to be at least 3-connected. 
%
Similarly, the results for the other cases are presented in the table.

@@

In Table~\ref{tab:smt-grph}, we present the running times for the sear
ch of
VTSs of sizes 2 to 10 that satisfy the variants.
and compare with our old encoding
({Old-e}) from~\cite{shukla}.
%
For the comparison between both encodings we have fixed the total
number of molecules to be $|M| = 2|N|$ for $ |N|> 2$ and
$|M| = 2|N| + 1$ for $|N| = 2$.
%
For each variant, we fix maximum number of parallel
edges to 2.
%
In the table we have shown comparison for specific connectivity, for
example variant A is checked against any graph with connectivity 2,
variant B with connectivity 3 similarly for the rest of the Variants.
%

@@

\textbf{{\qbftool} (QBF based synthesis tool):}

\input{synth-table}
We have implemented the encodings in a tool
called~\ourtool\footnote{{\url{https://github.com/arey0pushpa/pyZ3}}}.
%
The tool takes a partially defined VTS as input in a custom designed
input language.
%
The input is then converted to the constraints over VTS. 
%
The tool can not only synthesize the above-discussed queries, but also their
combinations.
%
For example, our tool can modify labels of nodes or edges while
learning activity functions.
%
Our tool is developed in C++ and uses~\zthree~\cite{z3} infrastructure for
processing formulas. 
%
Since some of the formulas involve alternation of quantifiers over
Boolean variables Z3 is not a suitable choice for those examples.
%
We translate the formulas created by Z3 tool into a standard
QDIMACS~\cite{qdimacs} format and use as an input for QBF solvers. 
%
We use~\depqbf~\cite{lonsing2010depqbf} for solving of QBF formulas. 
%
Our tool includes about 7000 lines of code.

We have applied~\ourtool~on six partially defined VTSs.
%
The results are presented in table~\ref{tab:qf-graph} for both the solvers
\depqbf and \zthree.
%
To use~\zthree, we remove \texttt{Connected} constraints, such that the queries becomes
quantifier-free.
%
% The experiments were conducted on a machine,
% with \ashu{?}MHz processor, \ashu{?}GB memory, and 900s timeout.
%
%The experiments were conducted with a 900s timeout.
%
The first four VTSs are synthetic but inspire from literature for
typical motifs in VTSs. 
The third VTS is a subgraph of the last VTS.
%
%
The fifth VTS is taken from~\cite{burri2004complete}.
%
The last VTS represent mammalian SNARE map created by studying the literature references.

>>>>

%We have developed four different tools (\sattool, \smttool, \qbftool, \ourtool) for implementation of proposed encodings.
%
We have developed the {\vtstool} tool to conjecture the properties of the VTS (explained in Section 6). 
%
%
The {\vtstool} tool includes about 10,000 lines of code.
%
The tool is divided into three different blocks,
%
checking existential condition, checking the universal condition and performing synthesis on the incomplete VTS. 
%
The existential condition is subdivided into two sub-block, using bounded model checking and using constraint solving. 
%
We will refer to existential subblocks as {\sattool} and {\smttool}, universal block as {\qbftool} and synthesis block as {\ourtool}.
%

\subsection{Implementation}
{\sattool} is based on the architecture of a Bounded model checker (CBMC) and uses an enumeration of paths for encoding the reachability problem.
%
{\smttool} is based on the SMT solver Z3 and uses the efficient encoding of finding a least fixed point to solve reachability problem. 
%
{\qbftool} is based on the QBF encoding of the universal condition and uses {\depqbf} solver to solve the generated QBF. 
%
The {\ourtool} is a synthesis tool for synthesizing incomplete information in the input VTS using {\qbftool}.
%
All the experiments were done on a machine with  Intel(R) Core(TM) i7-2600 CPU @ 3.40GHz processor and 16GB RAM with 30 min (1800 sec) timeout.

@@

The CBMC architecture consists of a C-language front-end (CFE) and a SAT Solver back-end (SAT). 
%
CBMC front end accepts ANSI-C programs with some special annotations (called assumes and assertions) to express constraints on the model and properties to be checked. 
%

In our experiments, the vesicle transport network was modelled as a non-deterministic C-program manipulating a graph with labelled edges with the fusion rules and steady state properties expressed as constraints using assumes. 
%		
The correspondence relationship between connectedness conditions
and guarantees for steady state were expressed as assertions to be checked. 
%		
Given an exploration depth provided by the user, which in our case corresponds to the size of the graph, CBMC verifies the properties by executing the following steps:
	\begin{itemize}
	\item Convert (using CFE) the model and the properties into a Boolean formula (verification condition) such that the property is true of all behaviors of the model up to the specified
depth iff the Boolean formula is valid.
	\item Check (using a SAT solver) validity of the verification condition. CBMC reports successful verification if the formula is found to be valid. If not, then CBMC produces a counterexample,
an assignment to the variables of the model, that is a witness to the violation of the property.
	\end{itemize}

@@

\textbf{{\smttool} (SMT based tool)}:
	We have implemented the encodings for each variants using the python interface of Z3 in a tool {\smttool}. 
%
Our tool allows the user to choose a model and the size
of the network besides other parameters like connectivity and number of parallel edges. 
%		
It uses Z3 Python interface to build the needed constraints and applies Z3 solver on the constraints to find a model (a satisfying assignment that respects the constraints). 
%
This tool also translates the satisfying model found by Z3 into
a VTS and presents a visual output to the user in form of annotated graph. 
%
We also visually report the dropped edges required to disconnect the graph, it gives information about the connectivity of the graph.
%

\textbf{{\qbftool} (QBF based universal condition):}
{\qbftool} is developed in C++ and uses~\zthree~\cite{z3} infrastructure and implement the encodings of the VTS in QBF form (presented in section 6.5). 
%
Since some of the formulas involve alternation of quantifiers over
Boolean variables Z3 is not a suitable choice for those examples.
%
We translate the formulas created by Z3 tool into a standard
QDIMACS~\cite{qdimacs} format and use as an input for QBF solvers. 
%
We use~\depqbf~\cite{lonsing2010depqbf} for solving the QBF formulas.

\textbf{{\ourtool} (QBF based synthesis tool):}

\input{synth-table}
We have implemented the synthesis  encodings in a tool
called~\ourtool\footnote{{\url{https://github.com/arey0pushpa/pyZ3}}}.
%
The tool takes a partially defined VTS as input in a custom designed
input language.
%
The input is then converted to the constraints over VTS using the {\qbftool} interface. 
%
The tool can not only synthesize the above-discussed queries, but also their
combinations.
%
For example, our tool can modify labels of nodes or edges while
learning activity functions.

\subsection{Results}

To illustrate usability of our tool in the last column of the Table~\ref{tab:smt-grph}, we present the minimum connectedness needed for the different variants after applying our tool for sizes from 2 to 10. 
%
We found no graph for the variant A with constraints Ann and Aen. 
%
Replacing constraint on the node of Ann with Anb (variant B)
does not affect the outcome. 
%
If we allow every present molecule to stay active
(Ann) but constraint the edge by a boolean function (Aeb) the resultant VTS has to be at least 3-connected. 
%
Similarly, the results for other cases are presented in the table.

@@

In Table~\ref{tab:smt-grph}, we present the running times for the search of
VTSs of sizes 2 to 10 that satisfy the variants.
and compare with the {\sattool}.
%
For the comparison between both encodings we have fixed the total
number of molecules to be $|M| = 2|N|$ for $ |N|> 2$ and
$|M| = 2|N| + 1$ for $|N| = 2$.
%
For each variant, we fix maximum number of parallel
edges to 2.
%
In the table we have shown comparison for specific connectivity, for
example variant A is checked against any graph with connectivity 2,
variant B with connectivity 3 similarly for the rest of the Variants.
%

@@

\input{qbf-table}
%\textbf{Universal Condition:}
Table~\ref{tab:qbf-graph} presents the result of the QBF tool.

%\textbf{Synthesis}:

We have applied~\ourtool~on six partially defined VTSs.
%
The results are presented in Fig~\ref{tab:synth-graph} for both the solvers
\depqbf and \zthree.
%
To use~\zthree, we remove \texttt{Connected} constraints, such that the queries becomes
quantifier-free.
%
% The experiments were conducted on a machine,
% with \ashu{?}MHz processor, \ashu{?}GB memory, and 900s timeout.
%
%The experiments were conducted with a 900s timeout.
%
The first four VTSs are synthetic but inspire from literature for
typical motifs in VTSs. 
The third VTS is a subgraph of the last VTS.
%
%
The fifth VTS is taken from~\cite{burri2004complete}.
%
The last VTS represent mammalian SNARE map created by studying the literature references.

### 2018-12-27 papers/journal/intro.tex
intro addition

%\textbf{Components of the vesicle transport network}:
A characteristic feature of eukaryotic cells is the presence of multiple internal membrane-bound organelles. 
These organelles exchange molecules amongst themselves in small membrane-bound packets called vesicles. 
%
The recognition of cargo molecules to be loaded onto these vesicles and the recognition of the correct target organelles for each vesicle is performed by an intricate network of molecules called the vesicle transport system~\cite{alberts2002molecular}. 
%
We call the set of internal organelles, together with the molecular routes that connect them through vesicle exchange, the vesicle transport network. 
%
Vesicle traffic underpins nearly every aspect of eukaryotic cellular physiology, including
human cellular physiology.
%
Understanding how this system function is, therefore,
one of the central challenges of cell biology.

\vspace{0.5cm}

\textbf{Components of the vesicle transport network}:
The vesicle transport network is used by cells to transport molecules across compartments.
% 
The mechanism of vesicle transport can roughly be broken down into two steps: budding of vesicles from the source compartment and their fusion with the target compartment.
%
Vesicle budding involves three steps:
\begin{enumerate}
	\item  \textbf{Sorting of cargo molecules by cytosolic adaptor molecules}. 
	Cargo adaptors bind to short linear stretches of amino acids in the cargo molecules called sorting signals and thereby sequester them for packaging. 
	%
	Different adaptor proteins act at different compartments, for eg., sec23/24 is required
	for packaging vesicles that form at the ER, the AP2 adaptor is used for packaging vesicles at the plasma membrane, etc. 
	%
	Adaptors are recruited to the source compartments through interactions with
	the lipids that form the source compartment's membrane, and interactions with Arf/Sar GTPases which signal the initiation of vesicle formation~\cite{paczkowski2015cargo}.

>>>>

%\textbf{Components of the vesicle transport network}:
A characteristic feature of eukaryotic cells is the presence of multiple internal membrane-bound organelles called compartments.
% 
These compartments exchange molecules amongst themselves in small membrane-bound packets called vesicles. 
%
The recognition of cargo molecules to be loaded onto these vesicles and the recognition of the correct target compartments for each vesicle is performed by an intricate network of molecules called the vesicle transport system~\cite{alberts2002molecular}. 
%
We call the set of internal compartments, together with the molecular routes that connect them through vesicle exchange, the vesicle transport network. 
%
Vesicle traffic underpins nearly every aspect of eukaryotic cellular physiology, including
human cellular physiology.
%
Understanding how this system function is, therefore,
one of the central challenges of cell biology.
%
The high complexity of the problems~\cite{mani2016wine, mani2016stacking} in VTS restricts the use of traditional methods like simulation and statistical reasoning. 
%The high complexity of problem restricts the use of traditional methods like simulation and statistical reasoning and the requirement of precise analysis makes formal methods an ideal candidate. 
The requirement of precise analysis makes formal methods ideal candidate in this case.

First, we will give an overview of the VTS mechanism and then will sketch out a few interesting problems that are key to the understanding of the VTS.
%
%\textbf{Components of the vesicle transport network}:
The vesicle transport network is used by cells to transport molecules across compartments.
% 
The mechanism of vesicle transport can roughly be broken down into two steps: budding of vesicles from the source compartment and their fusion with the target compartment.
%
Vesicle budding involves three steps:
\begin{enumerate}
	\item  \textbf{Sorting of cargo molecules by cytosolic adaptor molecules}. 
	Cargo adaptors bind to short linear stretches of amino acids in the cargo molecules called sorting signals and thereby sequester them for packaging. 
	%
	Different adaptor proteins act at different compartments, for eg., sec23/24 is required
	for packaging vesicles that form at the ER, the AP2 adaptor is used for packaging vesicles at the plasma membrane, etc. 
	%
	Adaptors are recruited to the source compartments through interactions with
	the lipids that form the source compartment's membrane, and interactions with Arf/Sar GTPases which signal the initiation of vesicle formation~\cite{paczkowski2015cargo}.

### 2018-12-27 papers/journal/synth-table.tex
intro addition

\begin{table}[!ht]
%\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
\def\arraystretch{1.6}
\caption{
{\bf SNARE regulation and graph connectedness.}}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
\bf{Sr.No} & 	{\multirow{2}{*}{\textbf{Regulation}}} & 
  {\multirow{2}{*}{\textbf{Existential}}} &
{\multirow{2}{*}{\textbf{Universal}}}  \\ 
%	\cline{2-3} &  \cline{Connect & Timing & \\ \hline
%\thickhline

1. & Boolean function & Boolean function & 2C \\ \hline
2. & None & Boolean function & 3C \\ \hline
3. & Boolean function & SNARE-SNARE inhibition & 4C \\ \hline
4. & None & SNARE-SNARE inhibition & No graphs found \\ \hline
5. & Boolean function & None & No graphs found \\ \hline
6. &  None & None & No graphs found \\ \hline

%6. & Arbitrary Boolean & Specificity matrix and every edge is distinct & N \& S = 3C\\ \hline

\end{tabular}
\label{tab:smt-grph}
%\end{adjustwidth}
\end{table}

\begin{table}[t]
	\centering
	% \def\arraystretch{1.6}
	\begin{tabular}{|c|c|c|c|}
		\hline
		{\multirow{2}{*}{\textbf{Variant}}}  &
		\multicolumn{2}{c|}{\textbf{Constraints}} &
		{\multirow{2}{*}{\textbf{Graph connectivity}}}
		% \multicolumn{2}{c|}{\textbf{Graph connectivity}}
		\\
		\cline{2-3}
		&  \textbf{Rest} & \textbf{Activity} &  % & \textbf{Sampling: Guarantee}
	\\ \hline

	A. & \multirow{5}{*}{
		\makecell{ \ref{eq:f0}--\ref{eq:fuse2},\\
			\ref{eq:reach1},\ref{eq:reach2},\\
			\ref{eq:drop1}--\ref{eq:drop4}}
	}
	& \ref{eq:ann},\ref{eq:aen} & No graph     \\\cline{3-4}
	B. & & \ref{eq:anb},\ref{eq:aen} & No graph     \\\cline{3-4}
	C. & & \ref{eq:ann},\ref{eq:aeb} & 3-connected  \\\cline{3-4}
	D. & & \ref{eq:anb},\ref{eq:aeb} & 2-connected  \\\cline{3-4}
	E. & & \ref{eq:ann},\ref{eq:aep} & No graph     \\\cline{3-4}
	F. & & \ref{eq:anb},\ref{eq:aep} & 4-connected  \\\hline

	% C_es :Self edges are allowed
	% C_ed : Every edge is distinct 
\end{tabular}
\caption{{\bf Activity regulation of molecules vs. graph connectivity.
}}
\label{tab:smt-grph}
\end{table}

@@

	% \caption{Run-times for searching for Z3 models (in secs). \#C  stands for minimum changes.
	% Time is reported in seconds.}
	\label{tab:qbf-graph}
\end{sidewaysfigure}

>>>>

@@

	% \caption{Run-times for searching for Z3 models (in secs). \#C  stands for minimum changes.
	% Time is reported in seconds.}
	\label{tab:synth-graph}
\end{sidewaysfigure}
